---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# sYNGEO_Func_Sync_V2

# Code index

1. 01_single_traits_interpolated_abundances_new_cwm.R
2. 02_sync_single_traits_groups_cwm_cwv_interpolated.R
3. 03_distance_connectivity.R
4. 04_statistical_analysis.R



# 01_single_traits_interpolated_abundances_new_cwm.R

Takes raw abundance data and traits for each species
cleans and interpolates abundances. 
Produces CWMs and CMVs (interpolated) for all traits and groups


### output_data from scripts 02 & 03 are saved into output_data/sync under .gitignore

# 02_sync_single_traits_groups_cwm_cwv_interpolated.R

calculates synchrony, functional distance and diversity for each trait in each region
calculates synchrony in environmental variables


## 03_distance_connectivity.R

calculates distances and connectivity variable
joins to all synchrony datasets
join on water course distance for testing

## 04_statistical_analysis.R

statistical model for drivers of synchrony
all figures




# Computation:

## CWM and Variance

```{r}
devtools::install_github("alaindanet/ecocom")
library(ecocom)

#vector of species abundance
abun <- rpois(n = 10, lambda = 3)
# vector of species trait
trait <- rnorm(n = 10, mean = 10, sd = 3)

# Compute the moments of trait distribution
calc_cw_moments(trait = trait, weight = abun)

# 
calc_cw_mean(trait = trait, weight = abun)
calc_cw_variance(trait = trait, weight = abun)
```

## Synchrony

Source the function at the beginning of your script:
```{r}
source("https://raw.githubusercontent.com/alaindanet/fishcom/master/R/synchrony.R")

sync_mat <- matrix(
  c(0, 0, 1, 1),
  nrow = 2,
  byrow = TRUE,
  dimnames = list(
    paste0("t", c(1, 2)),
    c("sp1", "sp2")
  )
)

# Complete synchrony
sync_mat
compute_synchrony(cov(sync_mat))

# Complete asynchrony
async_mat <- sync_mat
async_mat[, 1] <- rev(async_mat[, 1])
compute_synchrony(cov(async_mat))
```

# Temperature data

## Moving average over 12 months

See [doc/b-temperature.html](https://github.com/ksirving/sYNGEO_Func_Sync_V2/blob/main/doc/b-temperature.html)

Water temperature data had too much NA starting from 2008 to be useful.

- Below is the air temperature:
  - annual moving average: `annual_avg`
  - summer moving average (june-july-august for all site but the ones located in
    Australia: november-december-january): `summer_avg`


```{r}
library(tidyverse)
air_annual_and_summer_avg <- read_csv(
  here("input_data", "Env", "air_annual_and_summer_avg.csv")
)

head(air_annual_and_summer_avg)
```

