---
title: "Temperature data"
author: "Alain Danet"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::html_document2:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
## target knits Rmds in their own session, so load libraries here.
## source("here::here(packages.R"))
library(tidyverse)
library(lubridate)
library(here)
```

```{r load-targets, include=FALSE}
fish_ab <- read_csv(here("input_data", "Bio", "fish_ab.csv")) %>%
  janitor::clean_names()
awt <- read_csv(here("input_data", "Env", "awt.csv")) %>%
  filter(year > 1980)
no_filt_wt <- read_csv(here("input_data", "Env", "no_filtered_water_temperature.csv"))  %>%
  filter(siteid %in% unique(fish_ab$site_id)) %>%
  filter(year(date) >= 1990)
  
filt_wt <- read_csv(here("input_data", "Env", "filtered_water_temperature.csv"))  %>%
  filter(siteid %in% unique(fish_ab$site_id)) %>%
  filter(year(date) >= 1990)

mv_avg_roll <- read_csv(here("input_data/Env/mv_avg_roll_water_tmp.csv"))
air_mv_avg_roll <- read_csv(here("input_data/Env/mv_avg_roll_air_temp.csv")) %>%
  filter(
    year(date) >= (min(fish_ab$year) - 1) &
      year(date) <= (max(fish_ab$year) + 1)
  )
# Those sites do not have dates, month, or quarter filled
missing_site <- unique(fish_ab$site_id)[!unique(fish_ab$site_id) %in% unique(air_mv_avg_roll$siteid)]


# site location
load(here("input_data/Env/site_desc_loc.rda"))
```

## Raw water temperature data

- Confirmation that water temperature goes to maximum
  end of 2010:

```{r}
summary(no_filt_wt$date)
```

- We see a kind of a break between 1995 and 2000, the average temperature
  increase from 8.25 to 8.6 celsius degrees. It might explain why so much
  temperature are NA  

```{r}
no_filt_wt %>%
  ggplot(aes(y = tmp, x = date)) +
  geom_smooth()
```

## Filtered water temperature data 

```{r}
summary(filt_wt)
```

```{r}
filt_wt %>%
  group_by(date) %>%
  summarise(
    n = n(),
    n_na = sum(is.na(tmp))
    ) %>%
  ggplot(aes(y = n, x = date)) +
  geom_line() +
  labs(y = "Number of sites")
```

```{r}
ti <- no_filt_wt %>%
  left_join(
    filt_wt %>%
      select(-tmp) %>%
      mutate(excluded = FALSE),
    by = c("siteid", "date")
  ) %>%
mutate(excluded = ifelse(is.na(excluded), TRUE, FALSE))

```
```{r}
ti %>%
  group_by(date) %>%
  summarise(
    n = n(),
    n_na = sum(excluded)
    ) %>%
  ggplot(aes(y = n_na, x = date)) +
  geom_line() +
  labs(y = "Number of sites")
```





## Moving average dataset 

- Keep only the site that are in the synchrony dataset:

```{r}
awt_lg <- awt %>%
  filter(siteid %in% unique(fish_ab$site_id)) %>%
  pivot_longer(
    -c(siteid, year),
    names_to = "variable",
    values_to = "mv_avg") %>%
  mutate(variable = str_replace_all(variable,
      c("tmp_a_ana" = "air",
        "tmp_w_ama" = "water"
      )
    )
  )
```

```{r}
awt_lg_site_avg <- awt_lg %>%
  group_by(year, variable) %>%
  summarise(
    n_na =  sum(is.na(mv_avg)),
    perc_na = n_na / length(mv_avg),
    mv_avg = ifelse(
      n_na / length(mv_avg) < .20,
      mean(mv_avg, na.rm = TRUE), NA
      ),
    .groups = "drop"
  )
```

```{r}
awt_lg %>%
  ggplot(aes(x = year, y = mv_avg, color = variable)) +
  geom_smooth() +
  labs(x = "Year", y = "Temperature (°C)", title = "Moving average window")
```
There is no water temperature data after 2008...
It is may be better to work with air temperature since data goes from 2004 to
2013.


```{r}
awt_lg_site_avg %>%
  ggplot(aes(x = year, y = perc_na, color = variable)) +
  geom_line(size = 3) +
  labs(x = "Year", y = "Percentage of NA")
```

```{r}
fish_ab %>%
  group_by(year) %>%
  summarise(n_op = length(unique(site_id))) %>%
  ggplot(aes(y = n_op, x = year)) +
  geom_line(size = 3) +
  labs(x = "Year", y = "# of sampling events")
```

```{r}
sites_water_air_tmean_av <- awt_lg %>%
  pivot_wider(names_from = "variable", values_from = "mv_avg")
write_csv(
  sites_water_air_tmean_av,
  here("input_data", "Env", "sites_water_air_tmean_av.csv")
)
```

- Lots of NA after 2005: 

```{r}
mv_avg_roll %>%
  mutate(year = year(date)) %>%
  group_by(siteid, year) %>%
  summarise(
    avg = mean(mv_avg_12m, na.rm = TRUE),
    n_na = sum(is.na(mv_avg_12m))
    ) %>%
  ggplot(aes(x = year, y = n_na)) +
  geom_smooth()
```



## 

- For air temperature, we computed annual average based on 12 months moving
  windows, as well as summer average over three months (July (june-july-august)
  but winter (novembre-decembre-january) for Australia).

```{r}
air_annual_and_summer_avg <- air_mv_avg_roll %>%
  left_join(site_desc_loc %>%
    select(siteid, country),
  by = "siteid") %>%
  mutate(
    year = year(date),
    month = month(date)
    ) %>%
  group_by(siteid, year) %>%
  summarise(
    annual_avg = mean(mv_avg_12m),
    summer_avg = mean(mv_avg_3m[month == ifelse(country != "AUS", 7, 12)]),
    .groups = "drop"
  )

# Check
mask_mv_avg <- air_mv_avg_roll$siteid == "S10015" & year(air_mv_avg_roll$date) == 2004
mask_annual_summer_avg <- air_annual_and_summer_avg$siteid == "S10015" &
  air_annual_and_summer_avg$year == 2004
## Test annual avg:
stopifnot(mean(air_mv_avg_roll[mask_mv_avg, ][["mv_avg_12m"]]) ==
  air_annual_and_summer_avg[mask_annual_summer_avg, ][["annual_avg"]])
## Test summer avg:
stopifnot(air_mv_avg_roll[mask_mv_avg, ][["mv_avg_3m"]][7] ==
  air_annual_and_summer_avg[mask_annual_summer_avg, ][["summer_avg"]])
## All good

write_csv(
  air_annual_and_summer_avg,
  file = here("input_data", "Env", "air_annual_and_summer_avg.csv")
)
```




## Analysis

## Reproducibility

<details><summary>Reproducibility receipt</summary>

```{r}
## datetime
Sys.time()

## repository
if(requireNamespace('git2r', quietly = TRUE)) {
  git2r::repository()
} else {
  c(
    system2("git", args = c("log", "--name-status", "-1"), stdout = TRUE),
    system2("git", args = c("remote", "-v"), stdout = TRUE)
  )
}

## session info
sessionInfo()
```

</details>
